Examining Crowd Work and Gig Work Through The Historical Lens of Piecework

### 2017-05-09 chi/piecework/presentation/complexity.tex
checkin presentation

\onlyinsubfile{
\usepackage[citestyle=numeric,backend=bibtex]{biblatex}
\bibliography{references}
  % \beamertemplategridbackground[1mm]
}

\begin{document}

@@

\begin{frame}{Complexity}
  What do we mean when we talk about complexity?
  \begin{itemize}
    \item<1-> Can crowds help you write something?\\
    \scriptsize{
      \textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}
    }
    \item<2-> Can crowds critique novel designs?\\
    \scriptsize{
      \textcite{yuanAlmost,fuge2014analysis}
    }
    \item<3-> Can crowds create things from whole cloth?\\
    \scriptsize{
      \textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}
    }
  \end{itemize}
\end{frame}

\begin{frame}{What Does Crowdsourcing Say?}
\begin{columns}
  \begin{column}{0.6\textwidth}
    \begin{itemize}
      \item Build complexity into the process
      \begin{itemize}
        \item<1> Apply CS methods to people\\
        \scriptsize{
          \textcite{crowdForgeKittur}
        }
        \item<2> Humans as computational units\\
        \scriptsize{
          \textcite{Lasecki:2014:LSR:2661334.2661352}
        }
        \item<3> Crowdsourcing workflows as function state machines\\
        \scriptsize{\textcite{latoza2014microtask}
        }
      \end{itemize}
    \end{itemize}
  \end{column}

@@

\begin{frame}{What Does Piecework Say?}
    % \begin{itemize}
      % \item
      George Airy (astronomer) used a very similar approach\\
      \scriptsize{\textcite{grier2013computers}}
      % \begin{itemize}
      %   \item Break work up so that individual tasks are simple\\
      %   Make the process \textit{itself} compile the final product.
      % \end{itemize}
    % \end{itemize}
    % \vspace{0mm}
    \begin{columns}
    \begin{column}{0.6\textwidth}
      \begin{figure}
        \includegraphics[width=\textwidth]{figures/photo/Greenwich-Observatory-computing-staff-1902.jpg}
      \end{figure}
    \end{column}

    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item Employed computers
        \item 13--20 years old
        \item Overworked
        \item Underpaid
        \item Could be fired at will
      \end{itemize}
    \end{column}
    \end{columns}

@@

\begin{frame}{George Airy}
    Airy built complexity into the process, assigning \textit{human computers} 
    to compute, verify, and correct the right ascension and declination of stars.

@@

\begin{frame}{Cottage Industry}
  % First appearances of piecework:
% \visible<2->{farms}%
% \visible<3->{, textiles}%
% \visible<4->{, and matchsticks.}
  \begin{columns}
  \visible<1->{
    \begin{column}{0.33\textwidth}
    \centering
    Farms
    \begin{figure}
    \includegraphics[width=\textwidth]{figures/idk.png}
    \end{figure}
    \end{column}
    }

  \visible<2->{
  \begin{column}{0.33\textwidth}
  \centering
    Textiles
  \begin{figure}
  \includegraphics[width=\textwidth]{figures/pieceworkers.jpg}
  \end{figure}
  \end{column}
  }

  \visible<3->{
  \begin{column}{0.33\textwidth}
  \centering
    Matchsticks
  \begin{figure}
  \includegraphics[width=\textwidth]{figures/photo/Matchgirl_strikers.PNG}
  \end{figure}
  \end{column}
  }
  \end{columns}
\end{frame}

\begin{frame}{Planes, Trains, and Automobiles}
  \begin{columns}
  % \vspace{1mm}
  \only<1,4->{
  \begin{column}{0.33\textwidth}
  \centering
  Trains
  \begin{figure}
    \begin{overlayarea}{\textwidth}{4cm}
    \begin{minipage}[c][4cm]{\textwidth}
      \includegraphics[width=\textwidth]{figures/photo/train_workers.jpg}
    \end{minipage}
    \end{overlayarea}
    % \caption{i wanna fly the train}
    \end{figure}
\end{column}

\only<1>{
    \begin{column}{0.66\textwidth}
      \begin{itemize}
        \item ``Efficiency experts'' measured how long it would take to do various jobs\\
        \scriptsize{\textcite{doi:10.2307/1883615}}
        \item These measurements would be used to assign values for each specific task\\
        \scriptsize{\textcite{american1921problem}}
        \item Train engineers instituted ``The Fix'' to correct perceived unfairness\\
        \scriptsize{\textcite{roy1954efficiency}}
      \end{itemize}
    \end{column}
  }
}

  \only<2,4->{
    \only<2>{
      \begin{column}{0.33\textwidth}
      \begin{itemize}
        \item Fordism, Taylorism, and Scientific Management in full force
      \end{itemize}
      \end{column}
    }

  \begin{column}{0.33\textwidth}
    \centering
      Automobiles
      \begin{figure}
        \begin{overlayarea}{\textwidth}{4cm}
        \begin{minipage}[c][4cm]{\textwidth}
          \includegraphics[width=\textwidth]{figures/photo/ford_assembly_line.jpg}
        \end{minipage}
        \end{overlayarea}
      \end{figure}
  \end{column}

  \only<2>{
    \begin{column}{0.33\textwidth}
    \begin{itemize}
      \item \textit{Manufacturing} proved amenable
            to assembly line processes.
    \end{itemize}
    \end{column}
  }
}

  \only<3->{
    \only<3>{
      \begin{column}{0.66\textwidth}
      \begin{itemize}
        \item Men drafted during World War II
        \item Factories turned to a new workforce who had neither conventional training nor experience
        \item Training for individual tasks and correspondingly assigning workers 
      \end{itemize}
      \end{column}
    }
  \begin{column}{0.33\textwidth}
    \centering
      Planes
      \begin{figure}
        \begin{overlayarea}{\textwidth}{4cm}
        \begin{minipage}[c][4cm]{\textwidth}
            \includegraphics[width=\textwidth]{figures/photo/Rosie_the_Riveter_(Vultee)_DS.jpg}
      \end{minipage}
      \end{overlayarea}
      \end{figure}
  \end{column}
  }

  \end{columns}

\begin{frame}{Comparisons}
\begin{itemize}
  \item Limited array of tasks versus arbitrarily complex work
  \begin{itemize}
    \item \textit{Building} planes versus \textit{fixing} trains
  \end{itemize}
  \item \normalsize{Has technology changed this?}
    \begin{itemize}
      \item Technology makes complex tasks relatively trivial
      \item Measuring workers is easier than ever
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Complexity} % {Cab Drivers}
% \begin{columns}
  % \begin{column}{0.20\textwidth}
    % \begin{itemize}
    %   \item 
      % Cab drivers
    % \end{itemize}
  % \end{column}

  % \begin{column}{0.80\textwidth}
  \begin{figure}
  \includegraphics<1>[width=\textwidth]{figures/photo/knowledge-student.jpg}
  \only<2>{
  \begin{columns}
  \begin{column}{0.2\textwidth}
  \includegraphics[width=\textwidth]{figures/photo/knowledge-student.jpg}
  \end{column}
  \begin{column}{0.8\textwidth}
  \includegraphics[width=\textwidth]{figures/complexity/brain_scans.png}
  \end{column}
  \end{columns}
  }
  \includegraphics<3>[width=\textwidth]{figures/photo/gps-map.jpg}
  \end{figure}
  % \end{column}
% \end{columns}
\end{frame}

>>>>

\onlyinsubfile{
}
\begin{document}

@@

\begin{frame}{Complexity}
  What do we mean when we talk about complexity?
  % \begin{minipage}[t][25mm][t]{\textwidth}
  %   \begin{columns}
  %     \begin{column}[T]{0.8\textwidth}
        \begin{itemize}
          \item<1->
          Can crowds help you write something?\\
          \scriptsize{\textcite{bernsteinSoylent,Kim:2014:CSI:2556288.2556986,Nebeling:2016:WCW:2858036.2858169}}\normalsize{}
          \item<2-> Can crowds critique novel designs?\\
          \scriptsize{\textcite{yuanAlmost,fuge2014analysis}}\normalsize{}
          \item<3-> Can crowds create things from whole cloth?\\
          \scriptsize{\textcite{KimStoria,Kim2017,Hahn:2016:KAB:2858036.2858364,Lasecki:2014:LSR:2661334.2661352}}\normalsize{}
        \end{itemize}
      % \end{column}
      % \begin{column}[T]{0.2\textwidth}
      %   \begin{figure}
    % \centering
    % \vspace{0mm}

    % \visible<1->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}}%
    % \visible<3->{
    % \includegraphics<1->[width=.5\textwidth]{figures/complexity/paper_resized.png}
    %               % \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
    % % }

    % \includegraphics<2->[width=.5\textwidth]{figures/complexity/chair_resized.png}

    % \includegraphics[width=.15\textwidth]{figures/paper-unfolded.png}$\rightarrow$
    % \includegraphics[width=.15\textwidth]{figures/rhino.png}
    % \includegraphics[width=.15\textwidth]{figures/fish.png}
    % \includegraphics[width=.15\textwidth]{figures/origami-1b.png}

    % \vspace{4mm}
    % \includegraphics[width=\textwidth]{figures/hexagonflattened.png}
    % \vspace{5mm}

    % \visible<2->{
    % \includegraphics[width=\textwidth]{figures/complexity/decompose.png}
    % }

    % \visible<4->{
    %   \vspace{5mm}
    %   \includegraphics[width=\textwidth]{figures/networkTransparency.png}
    % }
  % \end{figure}
  %     \end{column}
  %   \end{columns}
  %   \end{minipage}
\end{frame}

\begin{frame}{What Does Crowdsourcing Say?}
\begin{columns}
  \begin{column}{0.6\textwidth}
    \begin{itemize}
      \item Build complexity into the process
      \begin{itemize}
        \item<1> Apply CS methods to people\\
        \scriptsize{\textcite{crowdForgeKittur}}\normalsize{}
        \item<2> Humans as computational units\\
        \scriptsize{\textcite{Lasecki:2014:LSR:2661334.2661352}}\normalsize{}
        \item<3> Crowdsourcing workflows as function state machines\\
        \scriptsize{\textcite{latoza2014microtask}
        }
      \end{itemize}
    \end{itemize}
  \end{column}

@@

\begin{frame}{What Does Piecework Say?}
      George Airy (astronomer) used a very similar approach\\
      \scriptsize{\textcite{grier2013computers}}\normalsize{}
    \begin{columns}
    \begin{column}{0.6\textwidth}
      \begin{figure}
        \includegraphics[width=\textwidth]{figures/photo/Greenwich-Observatory-computing-staff-1902.jpg}
      \end{figure}
    \end{column}

    \begin{column}{0.4\textwidth}
      \begin{itemize}
        \item Employed computers
        \item 13--20 years old
        \item no particularly strong background in mathematics
        \item A basic understanding of logarithms, algebra, etc\dots
      \end{itemize}
    \end{column}
    \end{columns}

@@

\begin{frame}{George Airy}
    Airy built complexity into the process, assigning \emph{human computers} 
    to calculate \& verify the \emph{right ascension} and \emph{declination} of stars.

@@

\subfile{piecework-simple}
\subfile{piecework-complex}

\begin{frame}{Comparisons}
\begin{itemize}
  \item Limited array of tasks versus arbitrarily complex work
  \begin{itemize}
    \item \emph{Building} planes versus \emph{fixing} trains
  \end{itemize}
  \item \normalsize{Has technology changed this?}
    \begin{itemize}
      \item Technology makes \emph{some} complex tasks relatively trivial
      \item Measuring workers is easier than ever
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Complexity} % {Cab Drivers}
  \begin{figure}
  \includegraphics<1>[width=\textwidth]{figures/photo/knowledge-student.jpg}
  \only<2>{
  \begin{columns}
  \begin{column}{0.2\textwidth}
  \includegraphics[width=\textwidth]{figures/photo/knowledge-student.jpg}
  \end{column}
  \begin{column}{0.8\textwidth}
  \includegraphics[width=\textwidth]{figures/complexity/brain_scans.png}
  \end{column}
  \end{columns}
  }
  \includegraphics<3>[width=\textwidth]{figures/photo/gps-map.jpg}
  \end{figure}
\end{frame}

### 2017-05-09 chi/piecework/presentation/introduction.tex
checkin presentation

\documentclass[presentation]{subfiles}

\onlyinsubfile{
  % \usepackage[citestyle=numeric,backend=bibtex]{biblatex}
  % \bibliography{~/Library/texmf/bibtex/bib/local/references}
  % \beamertemplategridbackground[1mm]
}

\begin{document}
% \section*{Introduction}

% \frame<1-3>[frame1]{
\begin{frame}<-3>[label=returnable]{\only<-3>{Before We Get Started\dots}\only<4->{Old Wine in New Bottles}}
      \begin{itemize}
        \item<1-> \alert{Crowd work}: digitally mediated \emph{information work}
        (for example, work done on Amazon Mechanical Turk, UpWork, or 99designs)\\
          \scriptsize{\textcite{crowdworkFuture}}
        \item<2-> \alert{Gig work}: digitally mediated (often \emph{physically embodied}) one--off jobs,
        such as
        {driving},
        {courier services},
        and {administrative support}\\
          \scriptsize{\textcite{friedman2014workers,Parigi:2016:GE:3026779.3013496}}
        \item<3-> \alert{On--demand work}: crowd work and gig work, collectively
        \item<5-> \alert{Piecework}: Payment for \emph{output} rather than for \emph{time}
      \end{itemize}
\end{frame}

\begin{frame}{Ongoing Threads in Crowdsourcing Research}
\begin{columns}
\begin{column}[T]{0.5\textwidth}
  \begin{enumerate}%[<+>]
    \item<2,4> \textbf{\large Complexity}\\
    \scriptsize{
    \textcite{suzukiAtelier,KimStoria,yuanAlmost,Yu2016b,Nebeling:2016:WCW:2858036.2858169,Hahn:2016:KAB:2858036.2858364}
    }
    \vspace{5mm}

    \item<3> \textbf{\large Decomposition}\\
    \scriptsize{
    \textcite{sensitiveTasks,LykourentzouPersonalityMatters,Law:2016:CKC:2858036.2858144,Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
    }
    \vspace{5mm}

    \item<5> \textbf{\large Relationships}\\
    \scriptsize{
      \textcite{turkopticon,storiesIraniSilberman,crowdcollab,takingAHITMcInnis,dynamo,uberAlgorithm}
    }

  \end{enumerate}
\end{column}

\begin{column}[T]{0.25\textwidth}
  \begin{figure}
    \centering
    % \vspace{-5mm}
    \vspace{0mm}

    \visible<2->{\includegraphics[width=.33\textwidth]{figures/complexity/geodesic.png}}%
    \visible<4->{\includegraphics[width=.33\textwidth]{figures/complexity/paper_resized.png}
                  \includegraphics[width=.33\textwidth]{figures/complexity/chair_resized.png}
    }

    \vspace{4mm}
    \includegraphics[width=\textwidth]{figures/hexagonflattened.png}
    \vspace{5mm}

    \visible<3->{
    \includegraphics[width=\textwidth]{figures/complexity/decompose.png}
    }

    \visible<5->{
      \vspace{5mm}
      \includegraphics[width=\textwidth]{figures/networkTransparency.png}
    }
  \end{figure}

\end{column}

\begin{column}[T]{0.25\textwidth}
\centering
\vspace{5mm}

\visible<2->{
  \small{Complexity}
  \vspace{13mm}
}

\small{Tasks}

\visible<3->{
  \vspace{11mm}
  \small{Decomposition}
}

\end{column}
\end{columns}
\end{frame}

@@

    What are the \alert{limits} of complexity in on--demand work?

    How can we \alert{reach} those limits?
\end{frame}

\begin{frame}{Thesis}
    These questions have all been asked before.

@@

\notinsubfile{
  \subfile{timeline.tex} % this takes a while to compile...
}

\begin{frame}{Introduction}
  We hope to provide:
      \begin{itemize}
        \item A useful ontological lens for making sense of on--demand work as a resurgence of \alert{piecework}.
        \item A method for making sense of contemporary phenomena through \alert{historical analysis}.
      \end{itemize}
\end{frame}

\begin{frame}{Comparative Historical Analysis}
\begin{itemize}
  \item Historical analysis isn't new
  \begin{itemize}
    \item In general\\
    \scriptsize{
      \textcite{rosenberg1994exploring,rosenberg1982inside}
    }
    \item In HCI\\
    \scriptsize{
      \textcite{Wyche2006,bodker1993historical}
    }
  \end{itemize}
  \item Still, it's an underutilized method
  \begin{itemize}
    \item Provide some basic framing for \emph{ostensibly} new phenomena
    \item \emph{Explicate} our theoretical grounding
    \item Flesh out \emph{differences} and their implications
  \end{itemize}
\end{itemize}
\end{frame}

>>>>

\documentclass[presentation]{subfiles}
\onlyinsubfile{}
\setbeamersize{description width=30mm}
\begin{document}
\begin{frame}<-3>[label=returnable]{\only<-3>{Before We Get Started\dots}\only<4->{Old Wine in New Bottles}}
      \begin{description}
        \item<1-> [Crowd work] Digitally mediated \alert{information work}
        (e.g. Amazon Mechanical Turk, UpWork)\\
          \scriptsize{\textcite{crowdworkFuture}}\normalsize{}
        \item<2-> [Gig work] Digitally mediated (often \alert{physically embodied}) one--off jobs,
        such as
        \emph{driving for hire},
        \emph{courier services},
        and \emph{administrative support}\\
          \scriptsize{\textcite{friedman2014workers,Parigi:2016:GE:3026779.3013496}}\normalsize{}
        \item<3-> [On--demand~work] Crowd work and gig work, collectively
        \item<5-> [Piecework] Payment for \emph{output} rather than for \emph{time}
      \end{description}
\end{frame}

\begin{frame}[standout]
On--demand work is a modern instantiation of a much older phenomenon
--- \alert{piecework}.

{\normalsize The historical arc of piecework
can shed light on persistent questions
in this ongoing phenomenon of on--demand work.}
\end{frame}

\begin{frame}[t]{What is piecework?}
\centering
  \emph{Payment for \textbf{output} rather than for \textbf{time}}
\vspace*{2mm}
\begin{columns}
  \begin{column}{0.33\textwidth}
    \centering
    Textiles
    \begin{figure}
    \begin{overlayarea}{\textwidth}{33mm}
    \begin{minipage}[c][33mm]{\textwidth}
      \includegraphics[width=\textwidth]{figures/pieceworkers.jpg}
    \end{minipage}
    \end{overlayarea}
    \end{figure}
  \end{column}
  \begin{column}{0.33\textwidth}
    \centering
    Automobiles
    \begin{figure}
    \begin{overlayarea}{\textwidth}{33mm}
    \begin{minipage}[c][33mm]{\textwidth}
      \includegraphics[width=\textwidth]{figures/photo/ford_assembly_line.jpg}
    \end{minipage}
    \end{overlayarea}
    \end{figure}
  \end{column}
  \begin{column}{0.33\textwidth}
    \centering
    Metalwork
    \begin{figure}
    \begin{overlayarea}{\textwidth}{33mm}
    \begin{minipage}[c][33mm]{\textwidth}
      \includegraphics[width=\textwidth]{figures/photo/Rosie_the_Riveter_(Vultee)_DS.jpg}
    \end{minipage}
    \end{overlayarea}
    \end{figure}
  \end{column}
\end{columns}

\visible<2->{
  \begin{columns}
  \centering
    \begin{column}{0.33\textwidth}
    \centering
      Crowd work
    \begin{figure}
      \begin{overlayarea}{\textwidth}{15mm}
      \begin{minipage}[c][15mm]{\textwidth}
        \includegraphics[width=\textwidth]{figures/amt.png}
      \end{minipage}
      \end{overlayarea}
    \end{figure}
    \end{column}
    \begin{column}{0.33\textwidth}
    \centering
      Gig Work
    \begin{figure}
    \begin{overlayarea}{\textwidth}{15mm}
      \begin{minipage}[c][15mm]{\textwidth}
        \includegraphics[width=\textwidth]{figures/uber.png}
      \end{minipage}
      \end{overlayarea}
    \end{figure}
    \end{column}
  \end{columns}
}

\end{frame}

% \begin{frame}{Why Piecework?}
% What makes on--demand work another instantiation of piecework?

% \ali{piecework }
% \end{frame}

@@

    What are the \alert{limits} of complexity in on--demand work?
\end{frame}

\begin{frame}{Thesis}
    This question --- and others like it --- has been asked before.

@@

\notinsubfile{
  \subfile{timeline.tex}
}

\begin{frame}{Introduction}
  We hope to provide:
      \begin{itemize}
        \item A useful ontological lens for making sense of on--demand work as a resurgence of \alert{piecework}
        \item A method for making sense of contemporary phenomena through \alert{historical analysis}
      \end{itemize}
\end{frame}

\begin{frame}{Comparative Historical Analysis}
\begin{itemize}
  \item Historical analysis isn't new
  \begin{itemize}
    \item In general\\
    \scriptsize{\textcite{rosenberg1994exploring,rosenberg1982inside}}\normalsize{}
    \item In HCI\\
    \scriptsize{\textcite{Wyche2006,bodker1993historical}}\normalsize{}
  \end{itemize}
  \item Still, it's an underutilized method
  \begin{itemize}
    \item Provide some basic framing for \emph{ostensibly} new phenomena
    \item \emph{Explicate} our theoretical grounding
    \item Flesh out \emph{differences} and their implications
  \end{itemize}
\end{itemize}
\end{frame}
